---
title: "Reproduce_conditional2"
author: "XC"
date: "27/09/2021"
output:
  pdf_document:
    fig_caption: yes
csl: apa.csl
Extension: raw_tex
---
### what is \newcommand ?
ref: https://stackoverflow.com/questions/41655383/r-markdown-similar-feature-to-newcommand-in-latex

\newcommand{\Deltab} {\Delta} 
\newcommand{\intd} {\textrm{d}}
\newcommand{\Bmat} {B}
\newcommand{\Cmat} {C}
\newcommand{\cmat} {c}
\newcommand{\Imat} {I}
\newcommand{\bvec} {b}
\newcommand{\svec} {s}
\newcommand{\uvec} {u}
\newcommand{\omegab} {\omega}
\newcommand{\s}{s}
\newcommand{\h}{h}
\renewcommand{\b}{b}
\newcommand{\e}{e}
\newcommand{\z}{z}
\renewcommand{\v}{v}
\renewcommand{\u}{u}
\newcommand{\w}{w}
\renewcommand{\d}{d}
\newcommand{\Z}{Z}
\newcommand{\x}{x}
\newcommand{\Y}{Y}
\newcommand{\Yvec} {Y}
\newcommand{\Zvec}{Z}
\newcommand{\epsilonb}{\varepsilon}
\newcommand{bI} {I}
\newcommand{\bB}{B}
\newcommand{\bbeta}{\beta}
\newcommand{\bzero}{0}
\newcommand{\bSigma}{\Sigma}
\newcommand{\E}{E}
\newcommand{\cov} {\mathrm{cov}}
\newcommand{\var}{\mathrm{var}}
\newcommand{\tr}{\mathrm{tr}}
\newcommand{\diag}{\mathrm{diag}}
\newcommand{\vect}{\mathrm{vec}}
\newcommand{\Gau}{\mathrm{Gau}}
\newcommand{\RR}{\mathbb{R}}
\newcommand{\T}{{ \mathrm{\scriptscriptstyle T} }}

\renewcommand{\figurename}{Fig.}


## Setting up

```{r, message=FALSE}

# INLA
#install.packages("INLA",repos=c(getOption("repos"),INLA="https://inla.r-inla-download.org/R/stable"), dep=TRUE)
library(INLA)
```
```{r}
# For core operation
library(Matrix)
library(dplyr)
library(tidyr)
```
```{r}
# for plotting and for arranging the figures into panels for publication
library(ggplot2)
library(gridExtra)
library(grid)
#install.packages("extrafont")
library(extrafont)

```
```{r}
# provides the data
#install.packages("maptools")
library(maptools)

#install.packages("mapproj")
library(mapproj)

library(RandomFields)
```
```{r}
# contains a handy routing for computing CRPSs, crps: Continuous Ranked Probability Score

#install.packages("verification")
library(verification)
```
```{r}
# for parallel operations we will be requiring `foreach` and `doParallel`
#install.packages("foreach")
library(foreach)
library(doParallel)
```

Example below consider four models that vary only through the interaction function $b_o(h)$. The models are

\begin{equation*}
\begin{array}{ll}
\textrm{Model 1 (independent Mat{\'e}rns):} &b_o(\h) \equiv 0, \\
\textrm{Model 2 (pointwise dependence):} &b_o(\h) \equiv A\delta(\h), \\
\textrm{Model 3 (diffused dependence):} & \textrm{Model 4 with~} \Deltab = 0 \\
\textrm{Model 4 (asymmetric dependence):} &b_o(\h) \equiv \left\{\begin{array}{ll} A\{1 - (\|\h - \Deltab\|/r)^2\}^2, & \| \h - \Deltab\| \le r \\ 0, & \textrm{otherwise}, \end{array} \right. 
\end{array}
\end{equation*}

where $\Deltab = (\Delta_1, \Delta_2)^\T$ is a shift-parameter vector that captures asymmetry, 
  $r$ is the aperture parameter, 
  and $A$ is a scaling parameter. 
  In Models 3 and 4, $b_o(\h)$ is a shifted bisquare function defined on $\mathbb{R}^2$. 
  The covariance functions $C_{11}(\cdot)$ and $C_{2|1}(\cdot)$ are Matérn covariance functions. 
  For each model we also consider a \emph{reversed} dependence, where we switch $Y_2$ and $Y_1$. This gives us a total of eight models to fit and compare.


First, set program options, indicating which parts of the program we want to run and which parts we want to skip

```{r}
### Model choice
model_names <- c("independent","pointwise","moving_average_delta0","moving_average")
img_path <- "../paper/art"                  ## Where to save the figures
show_figs <- 1                              ## Show the figures in document
print_figs <- 0                             ## Print figures to file (leave =0)
LK_analysis <- 0                            ## Carry out likelihood analysis
LOO_analysis <- 0                           ## Carry out LOO analysis
Shifted_Pars_estimation <- 0                ## Fit shifted parimonious Matern
RF_estimation <- 0                          ## Carry out LOO with RFields
useMPI <- 0                                 ## MPI backend available?
```







