---
title: "Reproduce conditional approach"
author: "XC"
date: "6/16/2021"
output:
  pdf_document:
    fig_caption: yes
csl: apa.csl
Extension: raw_tex

---

```{r load-packages, include=FALSE}
library(INLA)

library(dplyr)
library(tidyr)
library(Matrix)

library(ggplot2)
library(gridExtra)
library(grid)
library(extrafont)

library(maptools)
library(mapproj)
library(RandomFields)

library(verification)

library(foreach)
library(doParallel)

library(bicon)
```



\newcommand{\Deltab} {\Delta}
\newcommand{\intd} {\textrm{d}}
\newcommand{\Bmat} {B}
\newcommand{\Cmat} {C}
\newcommand{\cmat} {c}
\newcommand{\Imat} {I}
\newcommand{\bvec} {b}
\newcommand{\svec} {s}
\newcommand{\uvec} {u}
\newcommand{\omegab} {\omega}
\newcommand{\s}{s}
\newcommand{\h}{h}
\renewcommand{\b}{b}
\newcommand{\e}{e}
\newcommand{\z}{z}
\renewcommand{\v}{v}
\renewcommand{\u}{u}
\newcommand{\w}{w}
\renewcommand{\d}{d}
\newcommand{\Z}{Z}
\newcommand{\x}{x}
\newcommand{\Y}{Y}
\newcommand{\Yvec}{Y}
\newcommand{\Zvec}{Z}
\newcommand{\epsilonb}{\varepsilon}
\newcommand{\bI}{I}
\newcommand{\bB}{B}
\newcommand{\bbeta}{\beta}
\newcommand{\thetab}{\theta}
\newcommand{\bzero}{0}
\newcommand{\bSigma}{\Sigma}
\newcommand{\E}{E}
\newcommand{\cov}{\mathrm{cov}} 
\newcommand{\var}{\mathrm{var}}
\newcommand{\tr}{\mathrm{tr}}
\newcommand{\diag}{\mathrm{diag}}
\newcommand{\vect}{\mathrm{vec}}
\newcommand{\Gau}{\mathrm{Gau}}
\newcommand{\RR}{\mathbb{R}}
\newcommand{\T}{{ \mathrm{\scriptscriptstyle T} }}
\renewcommand{\figurename}{Fig.}

## Setting up

\begin{equation*}
\begin{array}{ll}
\textrm{Model 1 (independent Mat{\'e}rns):} &b_o(\h) \equiv 0, \\
\textrm{Model 2 (pointwise dependence):} &b_o(\h) \equiv A\delta(\h), \\
\textrm{Model 3 (diffused dependence):} & \textrm{Model 4 with~} \Deltab = 0 \\
\textrm{Model 4 (asymmetric dependence):} &b_o(\h) \equiv \left\{\begin{array}{ll} A\{1 - (\|\h - \Deltab\|/r)^2\}^2, & \| \h - \Deltab\| \le r \\ 0, & \textrm{otherwise}, \end{array} \right. 
\end{array}
\end{equation*}

where $\Deltab = (\Delta_1, \Delta_2)^\T$ is a shift-parameter vector that captures asymmetry,
$r$ is the aperture parameter,
and $A$ is a scaling parameter.

In Models 3 and 4, $b_o(\h)$ is a shifted bisquare function defined on $\mathbb{R}^2$.

The covariance functions $C_{11}(\cdot)$ and $C_{2|1}(\cdot)$ are Matérn covariance functions. 


For each model we also consider a \emph{reversed} dependence, where we switch $Y_2$ and $Y_1$. This gives us a total of eight models to fit and compare.


```{r}
### Model choice

model_names <- c("independent","pointwise","moving_average_delta0","moving_average")
image_path <- "../paper/art"
show_figs <- 1              ## show figs in document
print_figs <- 0             ## Print figures to file (leave =0)
LK_analysis <- 0            ## log-likelihood analysis 
LOO_analysis <- 0           ## LOO analysis 
Shifted_Pars_estimation <- 0  ## Fit shifted parimonious Matern
RF_estimation <- 0            ## Carry out LOO with RFields
useMPI <- 0                                 ## MPI backend available?
```



## The data
The data were made available through the package \texttt{RandomFields}.
We first load the data

```{r}
data(weather, package = "RandomFields")
weather <- weather %>% data.frame()
weather %>% head(4) %>% print()
```

The `weather` table contains four fields, with latitude, longitude, pressure forecasting errors, and temperature forecasting errors for December 13, 2003 at 4 p.m. in the North American Pacific Northwest.

Since pressure and temperature have different units, we find a scaling factor by taking the ratio of the sample variances of the two variates, and computing its square root.

We will use this factor to scale the pressure variable.

```{r}
p_scale <- var(weather$pressure) / var(weather$temperature) %>% 
  sqrt() %>%
  as.numeric()
```
  

From this data frame we extract $Z_1$ and $Z_2$ and concatenate them into one long vector $Z$ through a function `form_Z`.

The vectors $Z_1$ and $Z_2$ are inverted if the model being analysed is greater than 4 (reversed model).

We also define `m1` as the number of observations of $Y_1$, `m2` as the number of observations of $Y_2$ and `m` as the total number of observations.

```{r}
form_z <- function(model_num, scale = T){
  Z1 <- matrix(weather$temperature)
  Z2 <- matrix(weather$pressure)
  
  if(scale) Z2 <- Z2 / p_scale   # scale pressure
  
  if(model_num > 4) {
    temp <- Z1   # move temperature values out of Z1 name into temp 
    Z1 <- Z2     # pressure values go into name Z1
    Z2 <- temp   # move original temperature values into Z2
  }
  
  Z <- rbind(Z1, Z2)  # concatenate
}


## Number of observations
m1 <- nrow(weather)
m2 <- nrow(weather)
m = m1 + m2
I_m1 <- Diagonal(m1)
```


## Process Discretisation

We approximate the processes as a sum of elemental basis functions (tent functions) constructed on a triangulation.

The triangulation is formed using the mesher in the `INLA` package,
while we provide a tailored function in the package `bicon`, `initFEbasis`


?initFEbasis: 
initialise a finite element basis which initialises an object of class FEBasis which defines a set of elemental ‘tent’ basis functions over a pre-specified triangulation in 2-D



which takes information from the `INLA` mesher and casts it into a `Mesh` object

We provide several methods associated with the `Mesh` class which will be useful for plotting later on.

Importantly, the `Mesh` object also contains information on the areas of the elements in the Voronoi tesselation, which will be used to approximate the integrations.


```{r, warning=F, message=F}
## discretizing process Y1,Y2 using triangular grid

## constructing mesh
mesh <- inla.mesh.2d(loc = weather[c("lon", "lat")],
             cutoff = 0,        # minimum allowed dist between pts, two pts further apart at most of this value will be replaced by single point
             max.edge = 0.75,   # the largest allowed tri edge length
             offset = 4)        # automatic extension dist

mesh_locs <- mesh$loc[, 1:2]  # [1:2071, 1:2]


## compute distances as in Gneiting(2010) -- greate-circle distance

### Greate circle distance: shortest distance between two points on the surface of a sphere, measured along the surface of the sphere not interior

d <- RFearth2dist(as.matrix(mesh_locs)) # transform coods from earth (ellipsoid) to cartesian
D <- as.matrix(d)
Dvec <- as.double(c(D))  

```


### understandings 1
```{r}

str(mesh)
head(mesh$loc, 3)
tail(mesh$loc, 3)

d <- RFearth2dist(as.matrix(mesh_locs))
d_matrx <- as.matrix(d)
rm(d_matrx)
rm(d)

dim(D) #2071 2071
```

```{r}
length(Dvec) # 2071 * 2071 = 4289041
```


```{r}
## obseration locs distance in cartesian

Dobs <- as.matrix(RFearth2dist(as.matrix(weather[c("lon", "lat")])))

Dobs_vec <- c(Dobs)
```
#### understandings 2
```{r}

obs_locs <- weather[c("lon", "lat")] ## df 157 obs, 2 vars
str(obs_locs)

obs_locs <- as.matrix(weather[c("lon", "lat")]) # [1:157, 1:2]

length(Dobs_vec)  # 24649
```



```{r}
## Cast into custom Mesh object
## define a set of tent basis functions over a prespecified triangulation in 2D
Mesh <- initFEbasis(p = mesh_locs,
            t = mesh$graph$tv,
            K = mesh$graph$vv)

str(Mesh)
```




